Sampling Strategies in Decisions from Experience

### 2022-03-31 manuscript.Rmd
Update simulation study

abstract: "In decisions from experience, systematic choices between prospects are based on the assessment of differences in their latent properties. Yet, distinct properties can be used to distinguish among the prospects, and the choices may vary respectively. This paper presents a model that shows how the mind may rely on sampling strategies---i.e., the probability of sampling outcomes from different prospects in direct succession (switching probability)---to alter which differences in the latent properties evidence is accumulated for. Specifically, the model sequentially samples outcomes from two prospects and accumulates the results obtained from a sequence of ordinal comparisons between sample means, with switching probabilities determining the size of the samples underlying each comparison and thresholds determining how much comparisons a prospect must win to be chosen. In a computational analysis, the model is used to simulate the information-processing chain for choices between two prospects, while varying switching probabilities and thresholds. The analysis shows that the accumulation process provides evidence for differences in the expected values when low switching probabilities cause each comparison to be carried out on large samples. In turn, when high switching probabilities cause the accumulation process to unfold over comparisons based on small samples, evidence is accumulated for differences in the probability to produce a higher outcome most of the time. The respective choices for high switching probabilities reflect an as-if underweighting of rare outcomes that becomes more pronounced with increasing thresholds, demonstrating that the reliance on many small samples can explain the as-if underweighting of rare outcomes, while a large number of sampled outcomes is fully accounted for. The analysis also shows that the differences in the information-processing chain translate to systematic differences in the parameter estimitates of cumulative prospect theory's value and weighting function."

@@

Below, the key predictions concerning the effects of the model parameters---i.e., the switching probability $\psi$ and the threshold $\theta$---on the length of the information-processing chain and the final choice are outlined.
Moreover, previous research has shown that differences as to how information about the properties of prospects is processed can translate to characteristic shapes of CPT's value and weighting function [see @pachurHowTwainCan2017; @zilkerNonlinearProbabilityWeighting2021].
Respectively, this section also contains predictions about the potential relations between $\psi$ and $\theta$ and the CPT parameters.

@@

Now, since for low switching probabilities the model tends to choose the prospect possessing the larger EV, the parameters of the value and weighting function are predicted to correspond to a linear weighting of outcomes and probabilities.
Thus, the lower the switching probabilities, the more should the choices translate to parameter estimates that satisfy $\alpha = 1$, $\gamma = 1$, and $\delta = 1$.
However, for high switching probabilities, the model predicts systematic deviations from the solutions of EV maximization, which take the form of an as-if underweighting of rare outcomes:
More specifically, for $\psi = 1$, the model tends to choose the risky prospect, if the probability of its high-rank outcome satisfies $p_x(x_{high}) > .5$, irrespective of its EV.
Vice versa, the model tends to choose the safe outcome, if $p_x(x_{high}) < .5$ is satisfied.
Accordingly, for high switching probabilities, the weighting pattern should be such that the high-rank outcome is overweighted for $p_x(x_{high}) > .5$ and otherwise underweighted.
Note now that the possible weighting patterns for choices between at most two-outcome prospects readily reduces to an over- or underweighting of the high-rank outcome [@zilkerNonlinearProbabilityWeighting2021]:
$$
\begin{aligned}
  \pi_{high} &= w(p_x(x_{high})) \\
  \pi_{low} &= w(p_x(x_{low}) + p_x(x_{high})) - w(p_x(x_{high})) \\
  &= 1 - \pi_{high}
  \; .
\end{aligned}
\tag{15}
$$
That is, the images of the weighting function equal the decision weights assigned to the high-rank outcome of the risky prospect.
Hence, for high switching probabilities, the graph of the weighting function is predicted to run below the identity line for $p_x(x_{high}) < .5$---leading to an underweighting of small-probability outcomes---and above the identity line for $p_x(x_{high}) > .5$---leading to an overweighting of high-probability outcomes.
Accordingly, the graph of the weighting function should take a S-shape, which is given by $\gamma > 1$ and $\delta = 1$.
Yet, since the model tends to approximate the solutions of EV maximization the lower the switching probabilities are, the S-shape should be most pronounced for high switching probabilities.
Thus, $\gamma$ should increase with $\psi$.
Note further that the predicted shape of the weighting function for high switching probabilities reflects the round-wise strategy's sensitivity to differences in the outcome's probabilities.
That is, the slope of the weighting function increases around the midpoint of the probability scale and flattens to both ends.
reflecting that the round-wise strategy is more sensitive to the differences in the 
Accordingly, the round-wise strategy

Finally, an increase in the switching probability should lead to a more concave value function.
Similar to the slope of the weighting function,

```{r}
predictions <- tibble("CPT parameter" = 
                        c("$\\gamma$",
                          "$\\delta$",
                          "$\\alpha$"),
                      "Predictions for changes in $\\psi$" = 
                        c("$\\geq 1$, increases with $\\psi$",
                          "1, no effect of $\\psi$",
                          "$\\leq 1$, decreases with increasing $\\psi$"))
apa_table(predictions, caption = "Model Predictions", note = "The prediction for changes in $\\psi$ should become more robust for increasing values of $\\theta$.", escape = FALSE)
```

@@

Below

@@

@@

The test set of $60$ choice problems, each consisting of a safe prospect and a two-outcome prospect, was obtained by stratified sampling from an initial set of $10,000$ choice problems.
The stratification was used to assure that some of the choice problems contain a small (high) outcome of a probability smaller than $.20$.
The exact procedure was as follows:
For each of the $10.000$ problems in the initial set, three outcomes were randomly drawn from a uniform distribution over the interval $[0, 20]$, with the smallest and highest of these three outcomes being assigned to the risky prospect in order to omit the case of dominant prospects.
The probability for the smaller outcome of the risky prospect, $p_{small}$, was drawn from a uniform distribution over the interval $[.01, .99]$, with the probability of the higher outcome being set to $p_{high} = 1-p_{small}$.
To obtain the $60$ choice problems of the test set, the initial set was divided into three subsets, where each subset contained either all problems with $p_{high} \in (0,.2)$ or $p_{high} \in [.2,.8]$ or $p_{high} \in (.8,1)$.
From each of these subset, then, $20$ choice problems were randomly sampled.

@@

The evidence accumulation model presented in this paper was implemented as a computational model to simulate the sampling process and the final choice for each of the $60$ choice problems in the test set.
All of the following parameter values were combined with each other:

* **Switching probability**: 
The switching probability $s$, defined as the probability of sampling outcomes from different prospects in direct succession, was varied in the interval $[.1, 1]$ in increments of $.1$.

* **Decision variable**: The decision variable of a prospect either counts the absolute number of won comparisons or how much more comparisons were won relative to the other prospect.

* **Threshold**: 
The value of the threshold $\theta$, determining how much (more) comparisons a prospect must win, was varied in the interval $[1,5]$ in increments of 1.

\noindent
In sum, 100 (parameter combinations) x 60 (choices) x 100 (agents) = 600,000 sampling processes and choices were simulated.

### Modeling

@@

@@

##gamma
gamma <- cpt %>%
  filter(parameter == "gamma") %>%
  ggplot(aes(s, mean, color = s, shape = boundary)) +
  facet_wrap(~a, nrow = 1, labeller = labeller(a = as_labeller(label_theta, default = label_parsed))) +
  labs(x = element_blank(), #omit axis title
       y = expression(gamma),
       color = "Switching\nProbability",
       shape = "Boundary") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0,1,.5)) +
  geom_point() +
  geom_pointrange(aes(ymin=`2.5%`, ymax=`97.5%`)) + 
  scale_shape_manual(values = c(4, 16)) +
  scale_color_viridis(option = "viridis") +
  theme_apa()

##delta
delta <- cpt %>%
  filter(parameter == "delta") %>% 
  ggplot(aes(s, mean, color = s, shape = boundary)) +
  facet_wrap(~a, nrow = 1) + 
  labs(x = "Switching Probability",
       y = expression(delta),
       color = "Switching\nProbability",
       shape = "Boundary") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0,1,.5)) +
  geom_point() +
  geom_pointrange(aes(ymin=`2.5%`, ymax=`97.5%`)) + 
  scale_color_viridis(option = "viridis") +
  scale_shape_manual(values = c(4, 16)) +
  theme_apa() + 
  theme(strip.text.x = element_blank()) #omit facet labels

##merge figures
ggarrange(gamma, delta, ncol = 1, nrow = 2, common.legend = TRUE, legend = "right", labels = "AUTO") 
```

@@

##alpha
alpha <- cpt %>%
  filter(parameter == "alpha") %>% 
  ggplot(aes(s, mean, color = s, shape = boundary)) +
  facet_wrap(~a, nrow = 1, labeller = labeller(a = as_labeller(label_theta, default = label_parsed))) +
  labs(x = element_blank(),
       y = expression(alpha),
       color = "Switching\nProbability",
       shape = "Boundary") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0,1,.5)) +
  geom_point() +
  geom_pointrange(aes(ymin=`2.5%`, ymax=`97.5%`)) + 
  scale_color_viridis(option = "viridis") +
  scale_shape_manual(values = c(4, 16)) +
  theme_apa()

##rho
rho <- cpt %>%
  filter(parameter == "rho") %>% 
  ggplot(aes(s, mean, color = s, shape = boundary)) +
  facet_wrap(~a, nrow = 1) + 
  labs(x = "Switching Probability",
       y = expression(rho),
       color = "Switching\nProbability",
       shape = "Boundary") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1,.5)) +
  geom_point() +
  geom_pointrange(aes(ymin=`2.5%`, ymax=`97.5%`)) + 
  scale_color_viridis(option = "viridis") +
  scale_shape_manual(values = c(4, 16)) +
  theme_apa() + 
  theme(strip.text.x = element_blank())

##merge figure
ggarrange(alpha, rho, ncol = 1, nrow = 2, common.legend = TRUE, legend = "right", labels = "AUTO") 
```

@@

##merge figures
ggarrange(weight, value, ncol = 1, nrow = 2, common.legend = TRUE, legend = "right", labels = "AUTO") 
```

For each distinct parameter combination, we ran 20 chains of 40,000 iterations each, after a warm-up period of 1000 samples.
To reduce potential autocorrelation during the sampling process, we only kept every 20th sample (thinning).
The potential scale reduction factor $\hat{R}$ [@gelmanInferenceIterativeSimulation1992] was $\leq$ `r round(max(cpt$Rhat), 3)` for all parameters, indicating good convergence.
The minimum effective sample size was `r min(cpt$n.eff)`.

As expected, the scatter plot below shows an inverse relationship between switching probability and trial length.
I.e., the lower the switching probability, the larger become the sample sequences on which each comparison between prospects is based, which in turn leads to longer trials.  
This effect is particularly pronounced for low probabilities such that the increase in trial length accelerates as switching probability decreases.

The figures below display the estimates of the $\gamma$ and $\delta$ parameter of the probability weighting function [@prelecProbabilityWeightingFunction1998] fitted to DfE simulated for different parameter values of the generating model.
The estimates are plotted against the switching probability, where each panel represents a distinct boundary type and level, i.e., number of comparisons. 
Grey dots represent agent level estimates, colored dots represent the mean across all agent level estimates.

Most significantly, there is a strong relationship between the switching probability in the generating model and the $\gamma$ parameter.
I.e., large switching probabilities, which are indicative for small sample sequences, lead to larger estimates for $\gamma$. 
The resulting strong curvature leads to a compression of probabilities in the lower and upper range, reflecting underweighting of small probabilities and overweighting of large probabilities. 
This pattern is robust for varying degrees of the boundary level, except for $a = 1$. 
The latter deviation may be explained by the potential scale reduction factors for the respective parameter estimates, which indicate that the MCMC chains did not converge (see [Appendix 3][Appendix 3: Convergence Diagnostics for CPT Parameters]).

<!-- Plots Appendix -->

```{r A-data, include=FALSE}
#load choice data
choices_c <- read_csv("data/choices/choices_comprehensive.csv", col_types = cols_choices)
cpt_c <- read_csv("data/estimates/estimates_cpt_comprehensive.csv", col_types = cols_cpt)
```

```{r A-choice-rates, fig.height=7, fig.width=10, include=FALSE}
#prepare data
fr_rates_c <- choices_c %>%
  mutate(norm = case_when(ev_ratio > 1 ~ "A", ev_ratio < 1 ~ "B"),
         ev_ratio_exp = round(a_ev_exp/b_ev_exp, 2),
         norm_exp = case_when(ev_ratio_exp > 1 ~ "A", ev_ratio_exp < 1 ~ "B"))

fr_rates_c_l <- fr_rates_c %>%
  filter(!is.na(norm)) %>%
  group_by(s, boundary, a, rare, norm, choice) %>% 
  summarise(n = n()) %>%
  mutate(rate = round(n/sum(n), 2),
         type = case_when(norm == "A" & choice == "B" ~ "Safe",
                          norm == "B" & choice == "A" ~ "Risky")) %>%
  ungroup() %>%
  filter(!is.na(type)) %>%
  select(-c(norm, choice, n)) %>%
  mutate(norm_ev = "Latent")

fr_rates_c_s <- fr_rates_c %>%
  filter(!is.na(norm_exp)) %>%
  group_by(s, boundary, a, rare, norm_exp, choice) %>%
  summarise(n = n()) %>%
  mutate(rate = round(n/sum(n), 2),
         type = case_when(norm_exp == "A" & choice == "B" ~ "Safe",
                          norm_exp == "B" & choice == "A" ~ "Risky")) %>%
  ungroup() %>%
  filter(!is.na(type)) %>%
  select(-c(norm_exp, choice, n)) %>%
  mutate(norm_ev = "Sample")

fr_rates_c <- bind_rows(fr_rates_c_l, fr_rates_c_s) %>%
  mutate(rare = case_when(rare == "none" ~ "\\in \\[.2,.8\\]",
                          rare == "attractive" ~ "\\in (0,.2)",
                          rare == "unattractive" ~ "\\in (.8,1)"))

fr_rates_c %>%
  ggplot(aes(s, rate)) +
  facet_grid(rare~a, labeller = labeller(rare = as_labeller(label_rare, default = label_parsed), 
                                         a = as_labeller(label_theta, default = label_parsed))) +
  labs(x = "Switching Probability",
       y = "False Response Rate",
       color = "Normative\nExpected\nValue",
       shape = "False\nResponse\nType") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, .5)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, .5)) +
  geom_point(aes(color = norm_ev, shape = type)) +
  scale_shape_manual(values = c(4, 16)) +
  scale_color_viridis_d(option = "plasma", end = .7) +
  theme_apa()
```

```{r A-weighting-parameters, fig.height=7, fig.width=10, include=FALSE}
#prepare data
cpt_c <- cpt_c %>% mutate(boundary = if_else(boundary == "absolute", "Absolute", "Relative"))

##gamma
gamma_c <- cpt_c %>%
  filter(parameter == "gamma") %>%
  ggplot(aes(s, mean, color = s, shape = boundary)) +
  facet_wrap(~a, nrow = 1, labeller = labeller(a = as_labeller(label_theta, default = label_parsed))) +
  labs(x = element_blank(),
       y = expression(gamma),
       color = "Switching\nProbability",
       shape = "Boundary") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0,1,.5)) +
  geom_point() +
  geom_pointrange(aes(ymin=`2.5%`, ymax=`97.5%`)) + 
  scale_shape_manual(values = c(4, 16)) +
  scale_color_viridis(option = "plasma") +
  theme_apa()

##delta
delta_c <- cpt_c %>%
  filter(parameter == "delta") %>% 
  ggplot(aes(s, mean, color = s, shape = boundary)) +
  facet_wrap(~a, nrow = 1) + 
  labs(x = "Switching Probability",
       y = expression(delta),
       color = "Switching\nProbability",
       shape = "Boundary") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0,1,.5)) +
  geom_point() +
  geom_pointrange(aes(ymin=`2.5%`, ymax=`97.5%`)) + 
  scale_color_viridis(option = "plasma") +
  scale_shape_manual(values = c(4, 16)) +
  theme_apa() + 
  theme(strip.text.x = element_blank()) #omit facet labels

##merge figures
ggarrange(gamma_c, delta_c, ncol = 1, nrow = 2, common.legend = TRUE, legend = "right", labels = "AUTO") 
```

```{r A-value-parameters, fig.height=7, fig.width=10, include=FALSE}
#plot estimates

##alpha
alpha_c <- cpt_c %>%
  filter(parameter == "alpha") %>% 
  ggplot(aes(s, mean, color = s, shape = boundary)) +
  facet_wrap(~a, nrow = 1, labeller = labeller(a = as_labeller(label_theta, default = label_parsed))) +
  labs(x = element_blank(),
       y = expression(alpha),
       color = "Switching\nProbability",
       shape = "Boundary") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0,1,.5)) +
  geom_point() +
  geom_pointrange(aes(ymin=`2.5%`, ymax=`97.5%`)) + 
  scale_color_viridis(option = "plasma") +
  scale_shape_manual(values = c(4, 16)) +
  theme_apa()

##rho
rho_c <- cpt_c %>%
  filter(parameter == "rho") %>% 
  ggplot(aes(s, mean, color = s, shape = boundary)) +
  facet_wrap(~a, nrow = 1) + 
  labs(x = "Switching Probability",
       y = expression(rho),
       color = "Switching\nProbability",
       shape = "Boundary") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1,.5)) +
  geom_point() +
  geom_pointrange(aes(ymin=`2.5%`, ymax=`97.5%`)) + 
  scale_color_viridis(option = "plasma") +
  scale_shape_manual(values = c(4, 16)) +
  theme_apa() + 
  theme(strip.text.x = element_blank())

##merge figure
ggarrange(alpha_c, rho_c, ncol = 1, nrow = 2, common.legend = TRUE, legend = "right", labels = "AUTO") 
```

```{r A-cpt-graphs, fig.height=7, fig.width=10, include=FALSE}
#prepare data

###weighting function
wf_c <- cpt_c %>%
  select(s, boundary, a, parameter, mean) %>%
  pivot_wider(names_from = parameter, values_from = mean) %>%
  select(-c(alpha, rho)) %>%
  expand_grid(p = seq(0, 1, .05)) %>%
  mutate(w = round(  (delta * p^gamma)/ ((delta * p^gamma)+(1-p)^gamma), 2))

###value function
vf_c <- cpt_c %>%
  select(s, boundary, a, parameter, mean) %>%
  pivot_wider(names_from = parameter, values_from = mean) %>%
  select(-c(gamma, delta, rho)) %>%
  expand_grid(x = seq(0, 20, .5)) %>%  
  mutate(v = round(x^alpha, 2))

##split data to avoid passing two y-values to the same x-value in geom_line()
wf_c_a <- wf_c %>% filter(boundary == "Absolute")
wf_c_r <- wf_c %>% filter(boundary == "Relative")
vf_c_a <- vf_c %>% filter(boundary == "Absolute")
vf_c_r <- vf_c %>% filter(boundary == "Relative")

##weighting function
weight_c <- wf_c_a %>% 
  ggplot(aes(p, w, group = s, color = s)) +
  facet_wrap(~a, nrow = 1, labeller = labeller(a = as_labeller(label_theta, default = label_parsed))) + 
  labs(x = "Experienced Probability",
       y = "Decision Weight",
       color = "Switching\nProbability") +
  scale_x_continuous(breaks = seq(0, 1, .5)) +
  scale_y_continuous(breaks = seq(0, 1, .5)) +
  geom_line(size = .5) +
  geom_line(data = wf_c_r, size = .5) + 
  scale_color_viridis(option = "plasma") +
  theme_apa()

##value function
value_c <- vf_c_a %>% 
  ggplot(aes(x, v, group = s, color = s)) +
  facet_wrap(~a, nrow = 1)+ 
  labs(x = "Objective Outcome",
       y = "Subjective Value",
       color="Switching\nProbability") +
  scale_x_continuous(breaks = seq(0, 20, 10)) +
  scale_y_continuous(breaks = seq(0, 20, 10)) +
  geom_line(size = .5) +
  geom_line(data = vf_c_r, size = .5) + 
  scale_color_viridis(option = "plasma") +
  theme_apa() + 
  theme(strip.text.x = element_blank())

##merge figures
ggarrange(weight_c, value_c, ncol = 1, nrow = 2, common.legend = TRUE, legend = "right", labels = "AUTO") 
```

>>>>

abstract: "In decisions from experience, systematic choices between prospects are based on the assessment of differences in their latent properties. Yet, distinct properties can be used to distinguish among the prospects and the choices may vary respectively. This paper presents a model that shows how the mind may rely on sampling strategies---i.e., the probability of sampling outcomes from different prospects in direct succession (switching probability)---to alter which differences in the latent properties evidence is accumulated for. Specifically, the model sequentially samples outcomes from two prospects and accumulates the results obtained from a sequence of ordinal comparisons between sample means, with switching probabilities determining the size of the samples underlying each comparison and thresholds determining how much comparisons a prospect must win to be chosen. In a computational analysis, the model is used to simulate the information-processing chain for choices between two prospects, while varying switching probabilities and thresholds. The analysis shows that the accumulation process provides evidence for differences in the expected values when low switching probabilities cause each comparison to be carried out on large samples. In turn, when high switching probabilities cause the accumulation process to unfold over comparisons based on small samples, evidence is accumulated for differences in the probability to produce a higher outcome most of the time. The respective choices for high switching probabilities reflect an as-if underweighting of rare outcomes that becomes more pronounced with increasing thresholds, demonstrating that the reliance on many small samples can explain the as-if underweighting of rare outcomes, while a large number of sampled outcomes is fully accounted for. The analysis also shows that the differences in the information-processing chain translate to systematic differences in the parameter estimates of cumulative prospect theory's value and weighting function."

@@

Below, the key predictions concerning the effects of the model parameters---i.e., the switching probability $\psi$ and the threshold $\theta$---on the length of the information-processing chain and the final choice are outlined.
Moreover, previous research has shown that differences as to how information about the properties of prospects is processed can translate to characteristic shapes of CPT's value and weighting function [see @pachurHowTwainCan2017; @zilkerNonlinearProbabilityWeighting2021].
Respectively, this section also contains predictions about the potential relations between $\psi$ and $\theta$ and the CPT parameters.
A summary of the model predictions can be found in Table\ \@ref(tab:predictions).

@@

Now, since for low switching probabilities the model tends to choose the prospect possessing the larger EV, the parameters of the value and weighting function are predicted to correspond to a linear weighting of outcomes and probabilities.
Thus, the lower the switching probabilities, the more should the choices translate to parameter estimates that satisfy $\alpha = 1$, $\gamma = 1$, and $\delta = 1$.
However, for high switching probabilities, the model predicts systematic deviations from the solutions of EV maximization, which take the form of an as-if underweighting of rare outcomes:
More specifically, for $\psi = 1$, the model tends to choose the risky prospect, if the probability of its high-rank outcome satisfies $p_x(x_{high}) > .5$, irrespective of its EV.
Vice versa, the model tends to choose the safe outcome, if $p_x(x_{high}) < .5$ is satisfied.
Accordingly, for high switching probabilities, the weighting pattern should be such that the high-rank outcome is overweighted for $p_x(x_{high}) > .5$ and otherwise underweighted.
Note now that the possible weighting patterns for choices between at most two-outcome prospects readily reduces to an over- or underweighting of the high-rank outcome [@zilkerNonlinearProbabilityWeighting2021]:
$$
\begin{aligned}
  \pi_{high} &= w(p_x(x_{high})) \\
  \pi_{low} &= w(p_x(x_{low}) + p_x(x_{high})) - w(p_x(x_{high})) \\
  &= 1 - \pi_{high}
  \; .
\end{aligned}
\tag{15}
$$
That is, the images of the weighting function equal the decision weights assigned to the high-rank outcome of the risky prospect.
Hence, for high switching probabilities, the graph of the weighting function is predicted to run below the identity line for $p_x(x_{high}) < .5$---leading to an underweighting of small-probability outcomes---and above the identity line for $p_x(x_{high}) > .5$---leading to an overweighting of high-probability outcomes.
Accordingly, the graph of the weighting function should take a S-shape, which is given by $\gamma > 1$ and $\delta = 1$.
Yet, since the model tends to approximate the solutions of EV maximization the lower the switching probabilities are, the S-shape should be most pronounced for high switching probabilities.
Thus, $\gamma$ should increase with $\psi$.
Note further that the predicted shape of the weighting function for high switching probabilities reflects the round-wise strategy's high sensitivity to differences in the probability to produce the higher outcome most of the time.
That is, the slope of the weighting function increases around the midpoint of the probability scale.
The predicted flat ends of the graph, however, indicate that the round-wise strategy more or less ignores how much larger the probability to produce a higher outcome is.

Moreover, an increase in the switching probability should lead to a more concave value function, since all mean comparisons are ordinal and weighted equally in their contribution to the accumulated evidence.
Specifically, the model is such that the magnitudes of outcomes are ignored once a comparison is carried out.
Hence, for high switching probabilities, the choices depend on which prospect produces the higher outcome most of the time, irrespective of how much larger the outcomes are.
Therefore, $\alpha$ is predicted to decrease as $\psi$.
Finally, all predictions concerning the parameters of CPT's value and weighting function should become more robust with increases in $\theta$ as increasing thresholds lead to more systematic choices.

(ref:predictions) Summary of Model Predictions

```{r predictions}
predictions <- tibble("Measure" = 
                        c("Choices pattern",
                          "",
                          "$\\gamma$ (curvature of $w$)",
                          "$\\delta$ (elevation of $w$)",
                          "$\\alpha$ (concavity of $v$)"),
                      "Predictions for changes in $\\psi$" = 
                        c("More EV maximization for decreasing $\\psi$",
                          "More as-if underweighting of rare outcomes for increasing $\\psi$",
                          "$\\geq 1$, increases with $\\psi$",
                          "1, no effect of $\\psi$",
                          "$\\leq 1$, decreases with increasing $\\psi$"))
apa_table(predictions, caption = "(ref:predictions)", note = "The predictions for changes in $\\psi$ should become more robust for increasing values of $\\theta$ as choices become more systematic for high thresholds.", escape = FALSE)
```

@@

To substantiate and refine the predictions of the model, a computational implementation was used to simulate the sampling and accumulation process for a set of choices between a safe prospect and a two-outcome (risky) prospect.
The computational model was written in `R` [@rcoreteamLanguageEnvironmentStatistical2021] and is embodied in the GitHub repository accompanying this paper.
The simulation was conducted for varying combinations of switching probabilities and thresholds.
For each of these combinations, the simulated data was modeled with CPT's value and weighting function, using a Bayesian parameter estimation.
The simulated choices and estimated CPT parameters can be understood as theory-implied data and data models which remain to be tested against empirical data [see @guestHowComputationalModeling2021; @haslbeckModelingPsychopathologyData2021].

@@

Below, the data generation and data modeling procedure is briefly reviewed.

@@

A test set of $60$ choice problems was obtained by stratified sampling from an initial set of $10,000$ problems.
The stratification procedure was used to ensure that the test set contained choice problems with small-probability outcomes of different rank---i.e., small-probability outcomes that are either higher or lower than the EV of the two-outcome prospect---as well as choice problems without small-probability outcomes.
The exact procedure was as follows:
For each of the $10.000$ problems in the initial set, three outcomes were randomly drawn from a uniform distribution over the interval $[0, 20]$, with the smallest and highest of these three outcomes being assigned to the risky prospect to omit dominant prospects.
The probability for the lower outcome of the risky prospect, $p_x(x_{low})$, was drawn from a uniform distribution over the interval $(0, 1)$, with the probability of the higher outcome being set to $p_x(x_{high}) = 1-p_x(x_{low})$.
To obtain the $60$ choice problems of the test set, the initial set was divided into three subsets, where each subset contained either all problems with $p_x(x_{high}) \in (0,.2)$ or $p_x(x_{high}) \in [.2,.8]$ or $p_x(x_{high}) \in (.8,1)$.
From each of these subset, then, $20$ choice problems were randomly sampled.

@@

To simulate the sampling and accumulation processes, the following parameter values were used and combined with each other.
That is, each parameter combination was used to solve the test set.

* *Switching probability* $\psi$ : 
The switching probability was varied in the interval $[.1, 1]$ in increments of $.1$.

* *Threshold* $\theta$: 
The threshold was varied in the interval $[1,5]$ in increments of 1.

* *Decision variable*: The decision variable of a prospect either counts the number of won comparisons---implying an absolute threshold---or how much more comparisons were won relative to the other prospect---implying a relative threshold.

\noindent
Since the switching probability introduces some stochasticity into the information-processing chain, each of the 100 resulting parameter combinations was used 100 times---resembling 100 agents---to solve the test set. 
In sum, then 100 (parameter combinations) $\times$ 100 (agents) $\times$ 60 (choice problems)  = 600,000 sampling and accumulation processes were simulated. 
The respective dataset can be retrieved from the accompanying GitHub repository.

### Bayesian CPT Analysis of Choice Data

The choice data generated for each parameter combination was separately modeled in CPT, using Tversky and Kahneman's (1992) value function and Goldstein and Einhorn's (1987) weighting function.
To demonstrate that the distortions from linear probability weighting are not just due to the difference between the latent objective probabilities and the relative number of times the outcomes are sampled over the course of any accumulation process, the sampled relative frequencies are supplied as probability information to the weighting function.
To accommodate for possible random choice behavior that may be exerted by some parameter combinations of the data-generating accumulation model, the CPT model was amended with the logit choice rule
$$
p(P'^{safe}) = \frac{1} {1 + e^{-\rho(V_{safe}^{\frac{1}{\alpha}}-V_{risky}^{\frac{1}{\alpha}})}}
\; ,
\tag{16}
$$
where $p(P'^{safe})$ is the probability that the safe prospect is chosen over the risky prospect, $V_{safe}$ and $V_{risky}$ are the valuations of the safe and risky prospect that result from the parameter estimates of the value and weighting function, and $\rho \geq 0$ is a sensitivity parameter.
The sensitivity parameter governs how strongly the choice between prospects depends on the differences in the valuations of the prospects---i.e., the core CPT model.
Specifically, $\rho = 0$ implies no sensibility and therefore random choices.
In turn, for increasing values of $\rho$, the probability of choosing the prospect with the higher valuation increases.
Note further that the valuation of the prospects were rescaled by $\frac{1}{\alpha}$, as proposed by @stewartPsychologicalParametersHave2018, to guard against possible parameter interdependencies between $\alpha$ and $\rho$ [see @krefeld-schwalbStructuralParameterInterdependencies2021; @scheibehenneUsingBayesianHierarchical2015].
In sum, then, the model has four free parameters $\alpha, \gamma, \delta, \rho$.

Since for a given choice problem and parameter combination of the accumulation model there should be no systematic differences among the 100 iterations (agents), their data is combined and free parameters are estimated across them.
Such a grouped approach is expected to reveal the main effects of a given parameter combination of the accumulation model on the choice patterns and ignores unsystematic differences or noise that may be caused by the switching probability [see @nilssonHierarchicalBayesianParameter2011].
To facilitate comparisons across parameter combinations of the accumulation model, uninformative prior distributions were used throughout for all four parameters of the data model.
Specifically, the prior distribution $\alpha \sim U(0,1)$, $\gamma \sim U(0,2)$, $\delta \sim U(0,10)$, and $\rho \sim U(0,5)$ were used.

@@

To estimate the posterior distributions of the parameters, Markov Chain Monte Carlo sampling was used, where 20 chains of 40,000 samples each were run after a warm-up period of 1000 samples.
To reduce potential autocorrelation during the sampling process, only every 20th sample was kept (thinning), leaving a total of 40,000 samples across chains.
The minimum effective sample size was `r min(cpt$n.eff)`.
The potential scale reduction factor $\hat{R}$ [see @gelmanInferenceIterativeSimulation1992] was $\leq$ `r round(max(cpt$Rhat), 4)` for all parameters, indicating good convergence.

```{r}
choices %>%
  group_by(s, a) %>% 
  summarize(n_med = median(n_sample)) %>% 
  ggplot(aes(x = s, y = n_med, group = as.factor(a), color = as.factor(a))) + 
   labs(x = expression(psi),
       y = "Sampled Outcomes",
       color = expression(theta)) +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, .5)) +
  scale_y_continuous(limits = c(0, 150), breaks = seq(0, 150, 50)) +
  geom_point() +
  geom_line() + 
  scale_color_viridis_d(option = "viridis") +
  theme_apa()
```

@@

##gamma
gamma <- cpt %>%
  filter(parameter == "gamma") %>%
  filter(a != 1, boundary != "relative") %>% 
  ggplot(aes(s, mean, color = s, shape = boundary)) +
  facet_wrap(~a, nrow = 1, labeller = labeller(a = as_labeller(label_theta, default = label_parsed))) +
  labs(x = element_blank(), #omit axis title
       y = expression(gamma),
       color = "Switching\nProbability",
       shape = "Boundary") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0,1,.5)) +
  geom_point() +
  geom_pointrange(aes(ymin=`2.5%`, ymax=`97.5%`)) + 
  scale_shape_manual(values = c(4, 16)) +
  scale_color_viridis(option = "viridis") +
  theme_apa()

##delta
delta <- cpt %>%
  filter(parameter == "delta") %>% 
  filter(a != 1, boundary != "relative") %>% 
  ggplot(aes(s, mean, color = s, shape = boundary)) +
  facet_wrap(~a, nrow = 1) + 
  labs(x = "Switching Probability",
       y = expression(delta),
       color = "Switching\nProbability",
       shape = "Boundary") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0,1,.5)) +
  geom_point() +
  geom_pointrange(aes(ymin=`2.5%`, ymax=`97.5%`)) + 
  scale_color_viridis(option = "viridis") +
  scale_shape_manual(values = c(4, 16)) +
  theme_apa() + 
  theme(strip.text.x = element_blank()) #omit facet labels

##merge figures
ggarrange(gamma, delta, ncol = 1, nrow = 2, common.legend = TRUE, legend = "right") 
```

@@

##alpha
alpha <- cpt %>%
  filter(parameter == "alpha") %>% 
  filter(a != 1) %>% 
  ggplot(aes(s, mean, color = s, shape = boundary)) +
  facet_wrap(~a, nrow = 1, labeller = labeller(a = as_labeller(label_theta, default = label_parsed))) +
  labs(x = element_blank(),
       y = expression(alpha),
       color = "Switching\nProbability",
       shape = "Boundary") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0,1,.5)) +
  geom_point() +
  geom_pointrange(aes(ymin=`2.5%`, ymax=`97.5%`)) + 
  scale_color_viridis(option = "viridis") +
  scale_shape_manual(values = c(4, 16)) +
  theme_apa()

##rho
rho <- cpt %>%
  filter(parameter == "rho") %>% 
  filter(a != 1) %>% 
  ggplot(aes(s, mean, color = s, shape = boundary)) +
  facet_wrap(~a, nrow = 1) + 
  labs(x = "Switching Probability",
       y = expression(rho),
       color = "Switching\nProbability",
       shape = "Boundary") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1,.5)) +
  geom_point() +
  geom_pointrange(aes(ymin=`2.5%`, ymax=`97.5%`)) + 
  scale_color_viridis(option = "viridis") +
  scale_shape_manual(values = c(4, 16)) +
  theme_apa() + 
  theme(strip.text.x = element_blank())

##merge figure
ggarrange(alpha, rho, ncol = 1, nrow = 2, common.legend = TRUE, legend = "right") 
```

@@

##merge figures
ggarrange(weight, value, ncol = 1, nrow = 2, common.legend = TRUE, legend = "right") 
```
